---
title: "Untitled"
author: "Hagan"
date: "4/27/2022"
output:
  html_document:
    df_print: paged
---



```{r}
library(car)
library(semPlot)
library(psych)
library(knitr)
library(kableExtra)
library(MVN)
library(dplyr)
library(magrittr)
library(tidyr)
library(corrplot)
library(ggraph)

```


The Data Set

```{r}
data_fa = read.csv("D:/Hagan/PENS/Matakuliah/Semester 2/Statistika/UAS/data_fa percobaan.csv", header = TRUE, sep = ';')
data_fa$GENDER = as.factor(data_fa$GENDER)
data_fa$AGE = as.factor(data_fa$AGE)
data_fa$X2 = as.integer(data_fa$X2)
knitr::kable(head(data_fa),booktabs=TRUE,format="markdown")

```

```{r}
data_fa
```

```{r}
# create the working data frame by removing the ID variable
data_fa <- data_fa[,1:20]

```

1. Numeric and Graphical Description of the Data

1.1. Univariate descriptive statistics from the psych package.

```{r}
knitr::kable(describe(data_fa,type=2,fast=T),booktabs=TRUE,format="markdown")
```

1.2 Univariate Distribution Tests and Plots plus Evaluation of Multivariate
Normality

```{r}
x_vars <- data_fa
# use the mvn function for an extensive evaluation
# note that different kinds of tests can be specified with changes in the arguments
result <- mvn(data= x_vars, mvnTest="mardia", univariateTest="AD")
kable(result$univariateNormality, booktabs=TRUE)

```



```{r}
kable(result$multivariateNormality, booktabs=TRUE, format="markdown")

```



```{r}
#kable(mvn(data= x_vars,univariatePlot="histogram"), booktabs=TRUE, format="markdown")
```
1.2  Bivariate Characteristics of the data set
1.2.1

```{r}
scatterplotMatrix(data_fa,cex=.2,
smooth=list(col.smooth="red", spread=F, lwd.smooth=.3),
col="skyblue1",
regLine=list(lwd=.3,col="black"))
```

1.3 Covariances and Zero Order Correlations

```{r}
covmatrix1 <- round(cov(data_fa),digits=3)
knitr::kable(covmatrix1,booktabs=TRUE,format="markdown")
```
```{r}
mat1 <- cor(data_fa)
corrplot(mat1,type="upper",tl.pos="tp")
corrplot(mat1,add=T,type="lower", method="number",
col="black", diag=FALSE,tl.pos="n", cl.pos="n")
```

Using the lavaan package for CFA

2.1 Implement the CFA, First Model

```{r}
library(lavaan)
```

2.1.1 Define and fit the first model

```{r}
### CFA Model
wisc.model <- 'body=~ï..X1+X9+X12+X14
mind =~ X2+X11+X16
connection =~ X4+X5+X6
expertise =~ X7+X8+X10+X13+X17
meaning =~ X3+X15+X18+X19+X20'
```


```{r}
fit1 <- lavaan::cfa(wisc.model, data=data_fa,std.lv=TRUE)
summary(fit1, fit.measures=T,standardized=T)

```
2.1.2 Obtain coefficients


```{r}
# obtain only the coefficients
kable(coef(fit1), booktabs=TRUE, format="markdown")

```
2.1.3 Complete parameter listing

```{r}
parameterEstimates(fit1,standardized=T)

```


```{r}
parameterEstimates(fit1, standardized=TRUE) %>%
filter(op == "=~") %>%
select('Latent Factor'=lhs, Indicator=rhs, B=est, SE=se, Z=z, 'p-value'=pvalue, Beta=std.all) %>%
knitr::kable(digits = 3, booktabs=TRUE, format="markdown", caption="Factor Loadings")

```
2.1.4 Residuals correlation matrix

```{r}
cor_table <- residuals(fit1, type = "cor")$cov
#cor_table[upper.tri(cor_table)] <- # erase the upper triangle
#diag(cor_table) <- NA # erase the diagonal 0's
knitr::kable(cor_table, digits=3,format="markdown", booktabs=TRUE) # makes a nice table and rounds ever
```
2.1.5 Plot the residuals


```{r}
# extract the residuals from the fit1 model
# get rid of the duplicates and diagonal values
# create a vector for a
res1 <- residuals(fit1, type = "cor")$cov
res1[upper.tri(res1,diag=T)] <- NA
v1 <- as.vector(res1)
v2 <- v1[!is.na(v1)]
qqPlot(v2,id=F)
```

2.1.6 Modification Indices

```{r}
kable(modificationIndices(fit1, sort.=TRUE, minimum.value=3), booktabs=TRUE, format="markdown")

```
2.1.1 Path Diagram for the bifactor Model 1


```{r}
# Note that the base plot, including standardized path coefficients plots positive coefficients green
# and negative coefficients red. Red-green colorblindness issues anyone?
# I redrew it here to choose a blue and red. But all the coefficients in this example are
# positive,so they are shown with the skyblue.
# more challenging to use colors other than red and green. not in this doc
semPaths(fit1, residuals=F,sizeMan=7,"std",
posCol=c("skyblue4", "red"),
#edge.color="skyblue4",
edge.label.cex=1.2,layout="circle2")

```


```{r}
# or we could draw the paths in such a way to include the residuals:
#semPaths(fit1, sizeMan=7,"std",edge.color="skyblue4",edge.label.cex=1,layout="circle2")
# the base path diagram can be drawn much more simply:
#semPaths(fit1)
# or
semPaths(fit1,"std")
```

2.2 Generate a second model and compare

2.2.1 Add a path (Perf to comp) and Fit the second CFA model


```{r}
### CFA Model
wisc.model2 <- 'healthy=~ï..X1+X9+X12+X14+X2+X11+X16
connection =~ X4+X5+X6
expertise =~ X7+X8+X10+X13+X17
meaning =~ X3+X15+X18+X19+X20'

```


```{r}
fit2 <- lavaan::cfa(wisc.model2, data=data_fa,std.lv=TRUE)
summary(fit2, fit.measures=T,standardized=T)

```

2.2.2 Obtain coefficients

```{r}
knitr::kable(coef(fit2),booktabs=TRUE, format="markdown")
```


```{r}
parameterEstimates(fit2,standardized=TRUE)

```


```{r}
parameterEstimates(fit2, standardized=TRUE) %>%
filter(op == "=~") %>%
select('Latent Factor'=lhs, Indicator=rhs, B=est, SE=se, Z=z, 'p-value'=pvalue, Beta=std.all) %>%
knitr::kable(digits = 3, format="markdown", booktabs=TRUE, caption="Factor Loadings")
```
2.2.3 Residuals correlation matrix

```{r}
cor_table2 <- residuals(fit2, type = "cor")$cov
#cor_table[upper.tri(cor_table)] <- # erase the upper triangle
#diag(cor_table) <- NA # erase the diagonal 0's
knitr::kable(cor_table2, digits=3,format="markdown",booktabs=TRUE) # makes a nice table and rounds ever
```
2.2.4 Modification Indices for Model 2

```{r}
kable(modificationIndices(fit2, sort.=TRUE, minimum.value=3), booktabs=TRUE, format="markdown")

```

2.2.5 Path Diagram for Model 2
```{r}
# Note that the base plot, including standardized path coefficients plots positive coefficients green
# and negative coefficients red. Red-green colorblindness issues anyone?
# I redrew it here to choose a blue and red. But all the coefficients in this example are
# positive,so they are shown with the skyblue.
# more challenging to use colors other than red and green. not in this doc
semPaths(fit2, residuals=F,sizeMan=7,"std",
posCol=c("skyblue4", "red"),
#edge.color="skyblue4",
edge.label.cex=1.2,layout="circle2")
```


```{r}
# or we could draw the paths in such a way to include the residuals:
#semPaths(fit1, sizeMan=7,"std",edge.color="skyblue4",edge.label.cex=1,layout="circle2")
# the base path diagram can be drawn much more simply:
#semPaths(fit2)
# or
semPaths(fit2,"std")

```
2.3 Compare Model 1 and Model 2

```{r}
kable(anova(fit1,fit2), booktabs=TRUE, format="markdown")

```

SELESAI

addtional
2.4 An additional perspective on estimation and optimization

```{r}
### CFA Model
wisc.model <- 'body=~ï..X1+X9+X12+X14
mind =~ X2+X11+X16
connection =~ X4+X5+X6
expertise =~ X7+X8+X10+X13+X17
meaning =~ X3+X15+X18+X19+X20'
```


```{r}
fit1eqs <- lavaan::cfa(wisc.model, data=data_fa,std.lv=TRUE, mimic="EQS")
summary(fit1eqs, fit.measures=T,standardized=T)

```


```{r}

```